A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration.

### 2024-02-06 frontiers_submission/frontiers_submission.tex
manuscript revisions

\title[Meta-analysis of best management practices]{A meta-analysis of the impacts of best management practices on nonpoint source pollutants.}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

@@

In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41\% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}.
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.

@@

Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
\citet{linternBestManagementPractices2020} found 43\% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation.
However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices \citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, (2) failure to incorporate uncertainty in estimates \citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, and (3) assumption of static performance over time \citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.

One source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies \citep{linternBestManagementPractices2020}.
There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making \citep{agouridisLivestockGrazingManagement2005, barrettComparisonBMPPerformance2008, claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020, horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
These reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014, claryBMPPerformanceAnalysis2011, barrettComparisonBMPPerformance2008, grudzinskiDoesRiparianFencing2020}.
The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs \citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.
Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data \citep{barrettPerformanceComparisonStructural2005, liuReviewEffectivenessBest2017}.\\
A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs \citep{horvathEffectsRegionalClimate2023}.

\hypertarget{effect-size-calculations}{%
\subsection{Effect size calculations}\label{effect-size-calculations}}

@@

where \emph{x\textsubscript{control}} is the pre-treatment or control pollutant concentration and \emph{x\textsubscript{experiment}} is the pollutant concentration measured after the BMP intervention.
Several BMP data synthesis efforts have applied statistical summaries or regressions using BMP efficiency as the response variable of interest \citep{agouridisLivestockGrazingManagement2005, claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
There are several statistical shortcomings (distributional asymmetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes across multiple studies that are cause for concern for metrics estimated using this approach \citep{nuzzoPercentDifferencesAnother2018, coleStatisticsNotesWhat2017}.
\citet{barrettPerformanceComparisonStructural2005} demonstrated the use of effluent concentration directly as a response variable improved the ability to describe BMP performance.
More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
\citet{horvathEffectsRegionalClimate2023} used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (\textbf{hedges}).
\citet{grudzinskiDoesRiparianFencing2020} applied the log ratio of means (\emph{ROM}) to summarize performance of livestock BMPs.
\emph{ROM} quantifies the difference in means between the control and experimental group \citep{hedgesMetaanalysisResponseRatios1999}:

@@

The goals of this project were to (1) conduct a systematic review of BMP field studies; (2) develop estimates of FIB, nitrogen, phosphorus, and TSS removal using a meta-analytic approach; and (3) explore the effect of multiple moderators using multilevel random effects meta-analytic regression models.

@@

The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines \citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A\&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: ``fecal indicator bacteria'' OR ``\emph{E. coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR ``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (\emph{n} = 4) was assigned a subset of studies to evaluate if they should be included (Table \ref{tab:criteria}).
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.

@@

We used the ``rma.mv'' function in the \emph{metafor} R package to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
\emph{ROM} was used as the effect size which required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
Fifty-nine percent of 222 effect sizes were missing standard deviations required to estimate sampling variance.
Removal of studies due to missing variance information can introduce substantial bias \citep{kambachConsequencesMultipleImputation2020}.
Missing standard deviations were imputed using the pooled ratio of the mean effect size to coefficient of variation (CV) \citep{brackenStatisticalMethodsAnalysis1992}.
Sampling variance was estimated utilizing the average squared CV across all studies divided by sample size for each effect \citep{nakagawaRobustReadilyImplementable2023, doncasterCorrectionBiasMeta2018}:

@@

We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95\% CI {[}-3.23, 5.61{]}; TN: ROM = 0.31, 95\% CI {[}-1.15, 1.78{]}; DIN: ROM = -2.4, 95\% CI {[}-6.14, 1.59{]}; TP: ROM = -1.97, 95\% CI {[}-5.18, 1.23{]}; PO\textsubscript{4}: -0.05, 95\% CI {[}-2.82, 2.72{]}; TSS: ROM: 0.31, 95\% CI {[}-1.15, 1.78{]}; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
This approach repeatedly fits the selected model leaving out an individual value each time.
The overall effect and 95\% CI from each refit model is compared to the overall effect and 95\% CI of the model fit to the full dataset.
We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).

@@

There were a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
Our review was resticted to studies published after 1999.
The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS).
There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
This might suggest that spatially correlated predictors such as climate and soil variables would be difficult to model due to the poor spatial representation of the collected data.

@@

\caption{\label{tab:modelaic}Model selection AICc}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{AICc} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Candidate Models & FIB & TN & DIN & TP & PO\textsubscript4 & TSS\\
\midrule
$\sim$ Int & 208.9 & \textbf{37.7} & \textbf{112.6} & 96.7 & 42.1 & \textbf{103.9}\\
$\sim$ Int+Influent & 197.4 & 39.8 & 114.6 & \textbf{96.3} & \textbf{36.8} & 106.4\\
$\sim$ Int+AI & 209 & 40.3 & 114.8 & 98.2 & 45.2 & 106.5\\
$\sim$ Int+AI+Influent & 195 & 42.5 & 116.9 & 98.9 & 40.2 & 109.4\\
$\sim$ Int+BMP & 209.1 & 47.3 & 121 & 101.8 & 51.8 & 115.3\\
\addlinespace
$\sim$ Int+BMP+Influent & 196.7 & 48.9 & 124 & 102.3 & 45.6 & 120.4\\
$\sim$ Int+AI+BMP & 210.1 & 50.8 & 124.1 & 104.8 & 56.4 & 120.7\\
$\sim$ Int+AI+BMP+Influent & 195.6 & 52.9 & 127.3 & 105.6 & 50.8 & 126.6\\
$\sim$ Int+BMP\texttimes Influent & 201.3 & 67.5 & 133.7 & 121.5 & 62.9 & 174.1\\
$\sim$ Int+AI\texttimes BMP & 208.1 & 75.1 & 137.6 & 121 & 67.7 & 175.5\\
\addlinespace
$\sim$ Int+Influent+AI\texttimes BMP & \textbf{193.7} & 77.2 & 141.4 & 121.6 & 70.7 & 191.4\\
$\sim$ Int+AI+BMP\texttimes Influent & 202 & 83.2 & 145.9 & 127.3 & 72.9 & 191.9\\
$\sim$ Int+AI\texttimes BMP+BMP\texttimes Influent & 201.9 & 116.8 & 156.2 & 142.7 & 94.8 & 188.5\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}Int = intercept; AI = ariditiy index; BMP = BMP subcategory}\\
\end{tabular}
\end{table}

@@

We found 17 TP studies with 37 effect sizes and 9 PO\textsubscript{4} studies with 21 effect sizes for inclusion in regression models.
There was a significant overall reduction found for TP (ROM = 0.40, 95\% CI {[}0.03, 0.76{]}; BMP\textsubscript{eff} = 32.7\%, 95\% CI {[}3.4\%, 53.2\%{]}) but no evidence of negative or positive effect for PO\textsubscript{4} (ROM = -0.18, 95\% CI {[}-0.56, 0.19{]}; BMP\textsubscript{eff} = -20.1\%, 95\% CI {[}-75.3\%, 17.7\%{]}).
For both the TP and PO\textsubscript{4} models, heterogenity was high, with moderate to high within study variance and low to moderate between study variance (TP: \textit{I\textsuperscript{2}\textsubscript{total}} = 96.13, \textit{I\textsuperscript{2}\textsubscript{study}} = 32.15, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.99; PO\textsubscript{4}: \textit{I\textsuperscript{2}\textsubscript{total}} = 97.28, \textit{I\textsuperscript{2}\textsubscript{study}} = 33.36, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.92).
The best model for both parameters only included influent as a moderator (Table \ref{tab:modelaic}).
Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:phoseffect}; Table S5).

@@

Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}.

Meta-analysis indicated that BMPs showed significant overall reductions in FIB, TN, TP, and TSS concentrations when using ROM as an effect size.
We did not find strong evidence of overall leaching or reductions of DIN or PO\textsubscript{4} in BMP studies.
The results are in agreement to some extent with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.

compare overall effects to koch, barrett, clary

Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).
The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern)
Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)

>>>>

\title[Meta-analysis of best management practices]{A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration.}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

@@

In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41\% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}.
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.

@@

Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
\citet{linternBestManagementPractices2020} found 43\% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation.
However, field studies are much more likely to demonstrate mixed results including net releases (leaching) of pollutants under certain conditions \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of model parameters that represent management practices \citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, (2) the failure to incorporate the appropriate types of uncertainty into estimates \citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, and (3) the assumption of static performance over time \citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.

An underlying source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies \citep{linternBestManagementPractices2020}.
There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making \citep{agouridisLivestockGrazingManagement2005, barrettComparisonBMPPerformance2008, claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020, horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
These reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014, claryBMPPerformanceAnalysis2011, barrettComparisonBMPPerformance2008, grudzinskiDoesRiparianFencing2020}.
The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs \citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.

While it is assumed that site specific conditions are responsible for some of the heterogeneity in observed BMP performance, it is not clear how much of that variance is due to any one specific factor.
Influent concentration is likely to have some effect on certain types of structural BMPs.
\citet{barrettPerformanceComparisonStructural2005} demonstrated that percent pollutant reduction is often a function of influent quality.
Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations.
However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
Second, local climatic conditions can be expected to influence BMP performance.
BMPs in dry climates have been shown to be more likely to leach phosphorus than those in wetter climates \citep{horvathEffectsRegionalClimate2023}.
However, elucidating possible confounders such as climate and soil condition has been constrained by the lack of reported local condition data included in most BMP studies \citep{horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, eagleMetaAnalysisConstrained2017}.
The age and upkeep of BMPs is a third factor in BMP performance.
On one hand, the observed effects from BMPs are a function of various physical and biological processes that vary in the time required to produce desired reductions, especially as the spatial scale of the deployed project increases \citep{mealsLagTimeWater2010}.
These ``lag times'' between implementation and effect, which can be multiple years, have been shown to vary between parameter and BMP type \citep{mealsLagTimeWater2010}.
On the other hand the ability of BMPs to function effectively may also change over time.
There has not been overwhelming published evidence to demonstrate the change or lack of change in BMP performance over time \citep{liuReviewEffectivenessBest2017}.
Many of the papers and data available for assessing BMP performance are short term monitoring project, typically around 1 year or less in length \citep{liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014}, suggesting our ability to assess the long-term performance of BMPs is limited.

@@

where \emph{x\textsubscript{control}} is the pre-treatment or control pollutant concentration and \emph{x\textsubscript{experiment}} is the pollutant concentration measured after the BMP intervention.
Several BMP data synthesis efforts have applied statistical summaries or regressions using BMP efficiency as the response variable of interest \citep{agouridisLivestockGrazingManagement2005, claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
There are several statistical shortcomings (distributional asymmetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes across multiple studies that are cause for concern for metrics estimated using this approach \citep{nuzzoPercentDifferencesAnother2018, coleStatisticsNotesWhat2017}.
\citet{barrettPerformanceComparisonStructural2005} demonstrated the use of effluent concentration directly as a response variable improved the ability to describe BMP performance.
More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
\citet{horvathEffectsRegionalClimate2023} used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups \citep{hedgesStatisticalMethodsMetaanalysis1985}.
\citet{grudzinskiDoesRiparianFencing2020} applied the log ratio of means (\emph{ROM}) to summarize performance of livestock BMPs.
\emph{ROM} quantifies the difference in means between the control and experimental group \citep{hedgesMetaanalysisResponseRatios1999}:

@@

Building on previous work, the objectives of this paper are to (1) assess the general performance of BMPs in the published literature, and (2) identify relationships between BMP performance and potential effect size moderators.
To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify variables that explain heterogeneity in BMP performance.
Based on the existing literature we hypothesized that influent pollutant concentration, BMP type, and climate condition are influential in BMP performance and could be used to predict effluent concentration or percent reductions.

@@

The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines {[}\citet{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}; Figures S1-S2{]}.
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A\&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: ``fecal indicator bacteria'' OR ``\emph{E. coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR ``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (\emph{n} = 4) was assigned a subset of studies to evaluate if they should be included (Table \ref{tab:criteria}).
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.

@@

We used the ``rma.mv'' function in the \emph{metafor} R package to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
We fit separate models for FIB, total nitrogen (TN), dissolved inorganic nitrogen (DIN), total phosphorus (TP), orthophosphate (PO\textsubscript{4}), and total suspended sediment (TSS).
Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
\emph{ROM} was used as the effect size which required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
Fifty-nine percent of 222 effect sizes were missing standard deviations required to estimate sampling variance.
Removal of studies due to missing variance information can introduce substantial bias \citep{kambachConsequencesMultipleImputation2020}.
Missing standard deviations were imputed using the pooled ratio of the mean effect size to coefficient of variation (CV) \citep{brackenStatisticalMethodsAnalysis1992}.
Sampling variance was estimated utilizing the average squared CV across all studies divided by sample size for each effect \citep{nakagawaRobustReadilyImplementable2023, doncasterCorrectionBiasMeta2018}:

@@

We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95\% CI {[}-3.23, 5.61{]}; TN: ROM = 0.31, 95\% CI {[}-1.15, 1.78{]}; DIN: ROM = -2.4, 95\% CI {[}-6.14, 1.59{]}; TP: ROM = -1.97, 95\% CI {[}-5.18, 1.23{]}; PO\textsubscript{4}: -0.05, 95\% CI {[}-2.82, 2.72{]}; TSS: ROM: 0.31, 95\% CI {[}-1.15, 1.78{]}; Figure S3-S8); therefore, adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
This approach repeatedly fits the selected model leaving out an individual value each time.
The overall effect and 95\% CI from each refit model is compared to the overall effect and 95\% CI of the model fit to the full dataset.
We did not identify evidence of outliers or overly influential studies for any of our models (Figure S9-S14).

@@

Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
Our review was resticted to studies published after 1999.
The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS).
There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).

@@

\caption{\label{tab:modelaic}Model selection AICc.}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{AICc} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Candidate Models & FIB & TN & DIN & TP & PO\textsubscript4 & TSS\\
\midrule
$\sim$ Int & 208.9 & \textbf{37.7} & \textbf{112.6} & 96.7 & 42.1 & \textbf{103.9}\\
$\sim$ Int+Influent & 197.4 & 39.8 & 114.6 & \textbf{96.3} & \textbf{36.8} & 106.4\\
$\sim$ Int+AI & 209 & 40.3 & 114.8 & 98.2 & 45.2 & 106.5\\
$\sim$ Int+AI+Influent & 195 & 42.5 & 116.9 & 98.9 & 40.2 & 109.4\\
$\sim$ Int+BMP & 209.1 & 47.3 & 121 & 101.8 & 51.8 & 115.3\\
\addlinespace
$\sim$ Int+BMP+Influent & 196.7 & 48.9 & 124 & 102.3 & 45.6 & 120.4\\
$\sim$ Int+AI+BMP & 210.1 & 50.8 & 124.1 & 104.8 & 56.4 & 120.7\\
$\sim$ Int+AI+BMP+Influent & 195.6 & 52.9 & 127.3 & 105.6 & 50.8 & 126.6\\
$\sim$ Int+BMP\texttimes Influent & 201.3 & 67.5 & 133.7 & 121.5 & 62.9 & 174.1\\
$\sim$ Int+AI\texttimes BMP & 208.1 & 75.1 & 137.6 & 121 & 67.7 & 175.5\\
\addlinespace
$\sim$ Int+Influent+AI\texttimes BMP & \textbf{193.7} & 77.2 & 141.4 & 121.6 & 70.7 & 191.4\\
$\sim$ Int+AI+BMP\texttimes Influent & 202 & 83.2 & 145.9 & 127.3 & 72.9 & 191.9\\
$\sim$ Int+AI\texttimes BMP+BMP\texttimes Influent & 201.9 & 116.8 & 156.2 & 142.7 & 94.8 & 188.5\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}Int = intercept; AI = ariditiy index; BMP = BMP subcategory}\\
\end{tabular}
\end{table}

@@

We found 17 TP studies with 37 effect sizes and 9 PO\textsubscript{4} studies with 21 effect sizes for inclusion in regression models.
There was a significant overall reduction found for TP (ROM = 0.40, 95\% CI {[}0.03, 0.76{]}; BMP\textsubscript{eff} = 32.7\%, 95\% CI {[}3.4\%, 53.2\%{]}) but no evidence of negative or positive effect for PO\textsubscript{4} (ROM = -0.18, 95\% CI {[}-0.56, 0.19{]}; BMP\textsubscript{eff} = -20.1\%, 95\% CI {[}-75.3\%, 17.7\%{]}).
For both the TP and PO\textsubscript{4} models, heterogenity was high, with moderate to high within study variance and low to moderate between study variance (TP: \textit{I\textsuperscript{2}\textsubscript{total}} = 96.13, \textit{I\textsuperscript{2}\textsubscript{study}} = 32.15, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.99; PO\textsubscript{4}: \textit{I\textsuperscript{2}\textsubscript{total}} = 97.28, \textit{I\textsuperscript{2}\textsubscript{study}} = 33.36, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.92).
The best model for both parameters only included influent as a moderator (Table \ref{tab:modelaic}).
Moderators explained a relatively small amount of variance for both models (TP: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.12, PO\textsubscript{4}: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.35).
Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:phoseffect}; Table S5).

@@

\textbf{Not complete}

Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}.
Similar spatial disparities have been identified and discussed in \citet{kochNitrogenRemovalStormwater2014}, \citet{grudzinskiDoesRiparianFencing2020} and can be problematic for extrapolating results to other regions of interest.
Inconsistent spatial coverage presents a challenge for disentangling confounding spatially correlated predictors such as climate and soil due to poor representation within the dataset.
\citet{horvathEffectsRegionalClimate2023} found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}.
Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S. (Figure \ref{fig:aiindex}.
Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
Study scale, runoff sources and BMP types appeared well distributed, with the caveat that we are not aware of the actual distribution of these values in BMPs deployed across the country.
